---
title: "Example TESAsamSim Run and DLM fit"
author: "Brendan Connors and Carrie Holt. Updated by Luke Warkentin and Robyn Forrest"
date: "February 4, 2021"
output:
  html_document: default
  pdf_document: default
---

This document illustrates how to simulate some stock-recruitment data using the `TESAsamSim` R package and then fit a linear state-space model (aka Dynamic Linear model) with potentially time-varying parameters to it. The model is fit via Maximum likelihood followed by Kalman filtering and smoothing using a wrapper for the `dlm` R package (see [here](https://cran.r-project.org/web/packages/dlm/index.html)). The wrapper is based in part on code generously provided by Cody Szuwalski (NOAA). For this illustrative example the simulations are based on the population characteristics of Interior Fraser River coho salmon.

First, load packages, read in data and set up scenarios

```{r, echo=FALSE}

## Check if necessary packages are available and install if necessary
listOfPackages <- c("here", "parallel", "doParallel", "foreach", 
                    "tidyverse", "tictoc", "TESAsamSim")
newPackages <- listOfPackages[!(listOfPackages %in% 
                                  installed.packages()[ , "Package"])]
if (length(newPackages)) {
  install.packages(newPackages)
}
lapply(listOfPackages, require, character.only = TRUE)

# load wrapper function to fit dynamic linear models
source(here("dlm-wrapper.R"))

## Load relevant input data
# CU-specific parameters
cuPar <- read.csv(here("data", "IFCohoPars", "cohoCUpars.csv"),
                  stringsAsFactors=F)

# Stock-recruit and catch data that are used to populate the simulation priming/conditioning period
srDat <- read.csv(here("data", "IFCohoPars", "cohoRecDatTrim.csv"),
                  stringsAsFactors=F)

simPar <- read.csv(here("data", "IFCohoPars",
                         "cohoSimPar.csv"),
                       stringsAsFactors = F)

# Directory where simulated data will be saved
scenNames <- unique(simPar$scenario)
dirNames <- sapply(scenNames, function(x) paste(x, unique(simPar$species),
                                                sep = "_"))
# loop through each scenario and simulate data
for (i in 1:4){
 genericRecoverySim(simPar=simPar[i,], cuPar=cuPar,  srDat=srDat,
            variableCU=FALSE, ricPars=NULL, #ricPars,  #cuCustomCorrMat = cuCustomCorrMat,
            dirName="example", nTrials=4, makeSubDirs=FALSE, random=FALSE)}

```

Put all results in the same list then convert to a dataframe

```{r echo=F, , warning=FALSE}

# Read the four sim scenarios into a list
simData <- list()
simData[[1]] <- readRDS(here("outputs", "simData", "example",
                         "stationary_ref_CUsrDat.RData"))$srDatout
simData[[2]] <- readRDS(here("outputs", "simData", "example",
                         "increaseProd_ref_CUsrDat.RData"))$srDatout
simData[[3]] <- readRDS(here("outputs", "simData", "example",
                         "increaseCapacity_ref_CUsrDat.RData"))$srDatout
simData[[4]] <- readRDS(here("outputs", "simData", "example",
                         "increaseProd&Capacity_ref_CUsrDat.RData"))$srDatout
# filter for CU and rename cols
for(i in 1:4){
  simData[[i]] <- simData[[i]] %>% 
    dplyr::filter(CU == 1) %>% 
    mutate() %>% 
    select(-CU) %>% 
    rename(byr=year, spwn=spawners, rec=recruits, alpha_true=alpha, beta_true=beta) %>% 
    mutate(scen = scenNames[i], alpha=99., beta=99., alpha_se=99., beta_se=99.) %>% # cols for output
    select(scen, everything())  #reorder cols
}

#, alpha=9999., beta=9999., alpha_se=9999., beta_se=9999.

# convert to df with all scenarios
dlm_Out <- simData[[1]]
for(i in 2:4) dlm_Out <- rbind(dlm_Out, simData[[i]])

iter <- unique(simData[[i]]$iteration)
nsc <- length(scenNames)

# Figure out how to do this in purrr later. Need to look at fitDLM more closely
# Borrowing Brendan's code
for(j in 1:nsc){
  for(i in seq_along(iter)){
    
    dat <- dlm_Out %>% 
      dplyr::filter(scen == scenNames[j], iteration==i) %>% 
      select(-c(alpha,beta,alpha_se,beta_se)) #need to remove these for fitting
    
    dlm_model <- fitDLM(data = dat,
                        alpha_vary = TRUE,
                        beta_vary = TRUE)
    
    dlm_Out[which(dlm_Out$scen==scenNames[j] & dlm_Out$iteration==i),10:13] <- dlm_model$results[,10:13]
  }
}

head(dlm_Out)

# View(dlm_Out)
# plotDLM(dlm_model)


```



















